Internship Summary: Mid-Term Review

### 2020-07-30 RapportStage/Rapport.tex
HF part for spherium

The RHF wave function cannot model properly the physics of the system at large $R$ because the spatial orbitals are restricted to be the same, and, \textit{a fortiori}, it cannot represent two electrons on opposite side of the sphere. 
In the UHF formalism there is a critical value of $R$, called Coulson-Fischer point \cite{Coulson_1949}, at which a UHF solution appears and is lower in energy than the RHF one.
The UHF solution has broken symmetry as the two electrons tends to localize on opposite sides of the sphere. 
The spatial part of the UHF wave function is defined as
\begin{equation}\label{eq:UHF_WF}
	\Psi_{\text{UHF}}(\theta_1,\theta_2)=\phi_\alpha(\theta_1)\phi_\beta(\theta_2)
\end{equation}
where $\phi_\sigma(\theta)$ is the spatial orbital associated with the spin-$\sigma$ electrons ($\sigma = \alpha$ for spin-up electrons and $\sigma = \beta$ for spin-down electrons).
These one-electron orbitals are expanded in the basis of zonal spherical harmonics
\begin{equation}
	\phi_\sigma(\theta)=\sum_{\ell=0}^{\infty}C_{\sigma,\ell}Y_{\ell}(\theta)
\end{equation}
It is possible to obtain the formula for the ground state UHF energy in this basis set \cite{Loos_2009}:

\begin{equation}
E_{\text{UHF}} = T_{\alpha} + T_{\beta} + V
\label{eq:EUHF}
\end{equation}
with
\begin{gather}
	T_{\sigma} = \frac{1}{R^2} \sum_{\ell=0}^{\infty} C_{\sigma,\ell}^2 \, \ell(\ell+1)
	\\
	V = \frac{1}{R} \sum_{\ell_1,\ell_2,\ell_3,\ell_4=0}^{\infty} \sum_{L=0}^{\infty} 
	(-1)^{\ell_3+\ell_4} v^\alpha_{\ell_1,\ell_2,L} v^\beta_{\ell_3,\ell_4,L} 
\end{gather}
and
\begin{equation}
	v^\sigma_{\ell_1,\ell_2,L} 
	= \sqrt{(2\ell_1+1)(2\ell_2+1)} C_{\sigma,\ell_1}C_{\sigma,\ell_2}
	\begin{pmatrix}
		\ell_1 & \ell_2 & L 
		\\
		0 & 0 & 0
	\end{pmatrix}^2 
\end{equation}
is expressed in terms of the Wigner 3j-symbols \cite{AngularBook}.
\titou{STOPPED HERE.}

Equation \eqref{eq:EUHF} is obtained from the general form of the wave function \eqref{eq:UHF_WF}, but to be associated with a physical wave function the energy has to be stationary with respect to the coefficients. The general method is to use a self-consistent field procedure \cite{SzaboBook} to get the coefficients of the wave functions corresponding to physical solutions. We will work in a minimal basis, composed of $Y_{0}$ and $Y_{1}$, or equivalently, a s and p\textsubscript{z} orbitals, to illustrate the difference between the RHF and UHF solutions. In this basis there is a shortcut to find the stationary solutions. One can define the one-electron wave functions $\phi(\theta)$ using a mixing angle between the two basis functions. Hence we just need to minimize the energy with respect to the two mixing angles $\chi_\alpha$ and $\chi_\beta$.

\begin{equation}
	\phi_\sigma(\theta)= \cos(\chi_\sigma)Y_{0}(\theta) + \sin(\chi_\sigma)Y_{1}(\theta)
\end{equation}

>>>>

The RHF wave function cannot model properly the physics of the system at large $R$ because the spatial orbitals are restricted to be the same, and, \textit{a fortiori}, it cannot represent two electrons on opposite side of the sphere. 
In the UHF formalism there is a critical value of $R$, called Coulson-Fischer point \cite{Coulson_1949}, at which a UHF solution appears and is lower in energy than the RHF one.
The UHF solution has broken symmetry as the two electrons tends to localize on opposite sides of the sphere. 
The spatial part of the UHF wave function is defined as
\begin{equation}\label{eq:UHF_WF}
	\Psi_{\text{UHF}}(\theta_1,\theta_2)=\phi_\alpha(\theta_1)\phi_\beta(\theta_2)
\end{equation}
where $\phi_\sigma(\theta)$ is the spatial orbital associated with the spin-$\sigma$ electrons ($\sigma = \alpha$ for spin-up electrons and $\sigma = \beta$ for spin-down electrons).
These one-electron orbitals are expanded in the basis of zonal spherical harmonics
\begin{equation}
	\phi_\sigma(\theta)=\sum_{\ell=0}^{\infty}C_{\sigma,\ell}Y_{\ell}(\theta)
\end{equation}
It is possible to obtain the formula for the HF energy in this basis set \cite{Loos_2009}:
\begin{equation}
	E_{\text{HF}} = T_{\text{HF}} + V_{\text{HF}}
\label{eq:EHF}
\end{equation}
where the kinetic and potential energies are
\begin{align}
	T_{\text{HF}} & = \sum_{\sigma=\alpha,\beta} \frac{1}{R^2} \sum_{\ell=0}^{\infty} C_{\sigma,\ell}^2 \, \ell(\ell+1)
	&
	V_{\text{HF}} & = \frac{1}{R} \sum_{L=0}^{\infty} 
	v^\alpha_{L} v^\beta_{L} 
\end{align}
and
\begin{equation}
	v^\sigma_{L} 
	=  \sum_{\ell_1,\ell_2} \sqrt{(2\ell_1+1)(2\ell_2+1)} C_{\sigma,\ell_1}C_{\sigma,\ell_2}
	\begin{pmatrix}
		\ell_1 & \ell_2 & L 
		\\
		0 & 0 & 0
	\end{pmatrix}^2 
\end{equation}
is expressed in terms of the Wigner 3j-symbols \cite{AngularBook}.

The general method is to use a self-consistent field procedure as described in Ref.~\cite{SzaboBook} to get the coefficients of the wave functions corresponding to stationary solutions with respect to the coefficients $C_{\sigma,\ell}$, i.e.,
\begin{equation}
	\pdv{E_{\text{HF}}}{C_{\sigma,\ell}} = 0.
\end{equation}. 
Here, we work in a minimal basis, composed of $Y_{0}$ and $Y_{1}$, or equivalently, a s and p\textsubscript{z} orbital, to illustrate the difference between the RHF and UHF solutions. In this basis there is a shortcut to find the stationary solutions. One can define the one-electron wave functions 
\begin{equation}
	\phi_\sigma(\theta)= \cos(\chi_\sigma)Y_{0}(\theta) + \sin(\chi_\sigma)Y_{1}(\theta)
\end{equation}
using a mixing angle between the two basis functions for each spin manifold. 
Hence we just minimize the energy with respect to the two mixing angles $\chi_\alpha$ and $\chi_\beta$.
\titou{STOPPED HERE.}
